---
title: "p8105_hw6_nc2710"
author: "Nicole Comfort"
date: "11/25/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(ggridges)
library(readxl)
library(dplyr)
library(janitor)
library(patchwork)
library(viridis)
library(rvest)
library(purrr)
library(stringr) 

# set options for figures
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "95%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository. The code below loads the raw data (which we also saw on the previous homework):

```{r import homicide data}

homicide_data = read_csv(file = "./data/homicide-data.csv") # import dataset

```

The raw data is a dataframe regarding homicides from 28 different states that took place from Jan 1, 2007 to November 2015. The dataset is composed of `r nrow(homicide_data)` observations of `r ncol(homicide_data)` variables. The variables include: unique case ID, reported date, victim last name and first name, as well as demographic information of the victim (race, age, sex). There is also information on the location including city, state, latitude/longitude, and finally, the disposition of the case (e.g. whether or not the case was resolved).

The code chunk below creates a city_state variable (e.g. “Baltimore, MD”) and a binary variable indicating whether the homicide is solved. We also omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. We omit Tulsa, AL – this is a data entry mistake. The code modifies victim_race to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.

```{r data manipulation}

homicide_data = 
  homicide_data %>% 
  janitor::clean_names() %>% # clean names
  mutate(city_state = str_c(city, state, sep = ", ", collapse = NULL)) %>% # create a city_state variable (e.g. “Baltimore, MD”)
  ungroup() %>% 
  filter(city_state != "Tulsa, AL" | "Dallas, TX" | "Phoenix, AZ" | "Kansas City, MO") %>%  # omit cities
  mutate(is_solved = recode(disposition,    # create binary variable for whether case is solved
                       "Closed without arrest" = FALSE,
                       "Open/No arrest" = FALSE,
                       "Closed by arrest" = TRUE))

```


For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}

```

## Problem 2 

In this probelm, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

babysex: baby’s sex (male = 1, female = 2)
bhead: baby’s head circumference at birth (centimeters)
blength: baby’s length at birth (centimeteres)
bwt: baby’s birth weight (grams)
delwt: mother’s weight at delivery (pounds)
fincome: family monthly income (in hundreds, rounded)
frace: father’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
gaweeks: gestational age in weeks
malform: presence of malformations that could affect weight (0 = absent, 1 = present)
menarche: mother’s age at menarche (years)
mheigth: mother’s height (inches)
momage: mother’s age at delivery (years)
mrace: mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
parity: number of live births prior to this pregnancy
pnumlbw: previous number of low birth weight babies
pnumgsa: number of prior small for gestational age babies
ppbmi: mother’s pre-pregnancy BMI
ppwt: mother’s pre-pregnancy weight (pounds)
smoken: average number of cigarettes smoked per day during pregnancy
wtgain: mother’s weight gain during pregnancy (pounds)

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.


```{r}

```

